---
title: "p8105_hw2_sz3030"
author: "Shaocong Zhang"
date: "10/9/2021"
output: github_document
---
# Problem 1

## Read and clean the Mr. Trash Wheel sheet

```{r TrashWheelSheet}
library(readxl)
library(tidyverse)

trash.wheel = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                          sheet = "Mr. Trash Wheel",
                          range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls))
```

## Read and clean precipitation data for 2018 and 2019

```{r Precipitation}
preciptation_2018 = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                          sheet = "2018 Precipitation", 
                          skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month, total) %>%
  mutate(year = 2018)

preciptation_2019 = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                          sheet = "2019 Precipitation", 
                          skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month, total) %>%
  mutate(year = 2019)
```

# Combine precipitation datasets and convert month to a character variable

```{r Combine}
comb_precipitation = bind_rows(preciptation_2018, preciptation_2019) %>%
  mutate(month = month.name[month])
```

For the Mr.Trash Wheel sheet, there are `r nrow(trash.wheel)` observations and `r ncol(trash.wheel)` variables. The variables includes information on the dumpster number, date of collection, amount of total litter and litter type. What's more, the median number of sports balls in a dumpster in 2019 was `r trash.wheel %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`.

For the combined precipitation, there are `r nrow(comb_precipitation)` observations and `r ncol(comb_precipitation)` variables. The variables includes information about the year, month and its total precipitation. What's more, the total precipitation in 2018 is `r comb_precipitation %>% filter(year == 2018) %>% pull(total) %>% sum()`

# Problem 2

## First, clean the data in pols-month.csv.

```{r clean_pols}
pols_moth = read_csv("./fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>%
  mutate(month = month.name[month], 
    president = recode(prez_dem, "0" = "dem", "1" = "gop")) %>%
  select(-prez_dem, -prez_gop, -day)
```

## Second, clean the data in snp.csv using a similar process to the above.

```{r clean_snp}
snp = read_csv("./fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>%
  arrange(year, month) %>%
  mutate(month = month.name[month]) %>%
  select(year, month, close) 
```


## Third, tidy the unemployment data so that it can be merged with the previous datasets.

```{r tidy_unemployment}
unemployment = read_csv("./fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemployment") %>%
  mutate(month = recode(month, Jan = "January", Feb = "February",
                        Mar = "March", Apr = "April",
                        May = "May", Jun = "June",
                        Jul = "July", Aug = "August",
                        Sep = "September", Oct = "October",
                        Nov = "November", Dec = "December")) %>%
  rename(year = Year)
```

## Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r }
comb_prob2 = left_join(pols_moth, snp) %>%
  left_join(., unemployment)
```

The original pols-month.csv dataset is related to the number of national politicians who are democratic or republican at any given time. For pols_moth dataset, there are `r nrow(pols_moth)` observations and `r ncol(pols_moth)` variables. The variables are year(range from `r min(pols_moth$year)` to `r max(pols_moth$year)`), month(from January to December), the number of republican governors on the associated date, the number of national politicians who are democratic or republican(gov_gop, sen_gop, rep_gop, gov_dem, sen_dem and rep_dem) and the sitting president is democratic or republican.

The original snp.csv is  related to Standard & Poor’s stock market index. For the resulting snp dataset, there are `r nrow(snp)` observations and `r ncol(snp)` variables. The variables are year(range from `r min(snp$year)` to `r max(snp$year)`) and the closing values of the S&P stock index on the associated date.

The original unemployment.csv dataset is related to the percentage of unemployment of the associated year in months. For the resulting unemployment dataset, there are `r nrow(unemployment)` observations and `r ncol(unemployment)` variables. The variables are years(range from `r min(unemployment$year)` to `r max(unemployment$year)`), month(from January to December) and the unemployment percentage.

The merged dataset comb_prob2 has `r nrow(comb_prob2)` observations and `r ncol(comb_prob2)` variables. The dataset contains all of the variables above


# Problem 3

## Load and tidy the data

By looking through the dataset, we can see that the names of a categorical predictor and the case structure of string variables changed over time when comes to year 2012 and the former. For example, "BLACK NON HISPANIC" of Ethnicities in 2016 while "BLACK NON HISP" in 2012, name "Aleksandra" in 2013 contrast with "ALEKSANDER" in 2012.

```{r echo = FALSE}
popular_baby_names = read_csv("./Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
    childs_first_name = str_to_sentence(childs_first_name),
    ethnicity = str_to_sentence(ethnicity),
    ethnicity = recode(ethnicity, "Asian and paci" = "Asian and pacific islander",
                       "Black non hisp" = "Black non hispanic",
                       "White non hisp" = "White non hispanic")
  ) %>%
   distinct %>% 
  arrange(year_of_birth, ethnicity, rank)

#popular_baby_names[!duplicated(popular_baby_names),] # remove the duplicated rows

```

## Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year.

```{r}
popular_baby_names %>%
  filter(childs_first_name == "Olivia", gender == "FEMALE") %>%
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(names_from = "year_of_birth", values_from = "rank")
```

## Produce a similar table showing the most popular name among male children over time.

```{r}
popular_baby_names %>% 
  filter(gender == "MALE", rank == 1) %>% 
  select(year_of_birth, ethnicity, childs_first_name) %>% 
  pivot_wider(names_from = year_of_birth, values_from = childs_first_name)
```

## Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
popular_baby_names %>%
  filter(gender == "MALE", ethnicity == "White non hispanic", year_of_birth == 2016) %>%
  ggplot(aes(x = rank, y = count)) + geom_point()
```




