---
title: "p8105_hw2_sz3030"
author: "Shaocong Zhang"
date: "10/9/2021"
output: github_document
---
# Problem 1

## Read and clean the Mr. Trash Wheel sheet

```{r TrashWheelSheet}
library(readxl)
library(tidyverse)

trash.wheel = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                          sheet = "Mr. Trash Wheel",range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls))
```

## Read and clean precipitation data for 2018 and 2019

```{r Precipitation}
preciptation_2018 = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                          sheet = "2018 Precipitation", skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month, total) %>%
  mutate(year = 2018)

preciptation_2019 = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                          sheet = "2019 Precipitation", skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month, total) %>%
  mutate(year = 2019)
```

# Combine precipitation datasets and convert month to a character variable

```{r Combine}
comb_precipitation = bind_rows(preciptation_2018, preciptation_2019) %>%
  mutate(month = month.name[month])
```

For the Mr.Trash Wheel sheet, there are `r nrow(trash.wheel)` observations and `r ncol(trash.wheel)` variables. The variables includes information on the dumpster number, date of collection, amount of total litter and litter type. What's more, the median number of sports balls in a dumpster in 2019 was `r trash.wheel %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`.

For the combined precipitation, there are `r nrow(comb_precipitation)` observations and `r ncol(comb_precipitation)` variables. The variables includes information about the year, month and its total precipitation. What's more, the total precipitation in 2018 is `r comb_precipitation %>% filter(year == 2018) %>% pull(total) %>% sum()`

# Problem 2

## First, clean the data in pols-month.csv.

```{r clean_pols}
pols_moth = read_csv("./fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>%
  mutate(month = month.name[month], 
    president = recode(prez_dem, "0" = "dem", "1" = "gop")) %>%
  select(-prez_dem, -prez_gop, -day)
```

## Second, clean the data in snp.csv using a similar process to the above.

```{r clean_snp}
snp = read.csv("./fivethirtyeight_datasets/snp.csv") %>%
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>%
  arrange(year, month) %>%
  mutate(month = month.name[month]) %>%
  select(year, month, close) 
```


## Third, tidy the unemployment data so that it can be merged with the previous datasets.

```{r tidy_unemployment}

```

